### 2019-12-20 README.Rmd
updated docs

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```

# msBalance

## Overview

msBalance is an R project for analysis of balance responses in people with and without multiple sclerosis (MS).
The source data were created by Jessie Huisinga, Ph.D., and collected by Daniel Peterson, Ph.D.
The analysis code was primarily developed by J. Lucas McKay, Ph.D., M.S.C.R.

## Getting and installing msBalance

### What to read and edit
1. This document explains how to run the code. It is an html version of `README.Rmd` that is "knit," which you can think of as sort-of interpreted or compiled.
1. The document `R/msBalance.html` contains the output text and figures. It is the knit/interpreted version of `R/msBalance.Rmd`

### How to run the code
1. Alter `paths.m` in the main project directory to point to the correct data directory on your local machine.
1. Run the Matlab code to load .mat files and populate patient-level data
    * Run `m/loadData.m` to load source files and populate the `csv/` directory.
    * A future version of this could call the Matlab script from R if useful.
1. "knit" `R/msBalance.Rmd` to produce `R/msBalance.html`.

### Notes related to the fact that there is a mixture of Matlab and R code
1. The source files are in `.mat` format.
1. There is a mixture of matlab preprocessing code (files in `m/`) into `.csv` files that R can then read (files in `R/`).
1. A directory `csv/` is created to hold these files. The contents of this directory are not automatically loaded to GitHub.

### How to download from GitHub for development
1. Use RStudio to create a local copy of the project
    1. In RStudio, start a new Project: `File > New Project > Version Control > Git`
    1. In the “repository URL” paste: https://github.com/jlucasmckay/msBalance
    1. On OS X, you can make this under `tmp` - this is a temporary file that can be deleted later
    1. If there are files that need to be added to the remote, they can be placed in this folder or dragged / dropped into the GitHub interface.
1. Path locations that need to be read by Matlab and R can be specified in `msBalance/paths.m`.
1. Path locations might need to be modified in `msBalance/m/loadData.m`.
1. The main analysis is in `msBalance/R/msBalance.Rmd`
1. Stage changed files, commit, and push changes with the git panel.