### 2020-01-16 README.Rmd
reorganized manuscript

---
title: "Balance responses in MS"
author: J.L. McKay
date: "`r format(Sys.time(), '%d %B %Y')`"
output: github_document
always_allow_html: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      comment = "#>")
```

# Overview
`msBalance` is an R project for analysis of balance responses in people with and without multiple sclerosis (MS).
The source data were created by Jessie Huisinga, Ph.D., and collected by Daniel Peterson, Ph.D.
The analysis code was primarily developed by J. Lucas McKay, Ph.D., M.S.C.R.

# Use

## What to read and edit
1. The main document is `R/msBalance.Rmd`, which can be knit/interpreted into `R/msBalance.html`, which contains the main text and figures.
1. This document explains how to run the code.

## How to run the code
1. Alter `paths.m` in the main project directory so that: 
    1. `dataDirectory` points at the correct data directory on your local machine. This should contain the files `AllConsolidatedData_CS.mat` and `AllConsolidatedData_MS.mat`.
    1. `matlabCommand` points at the appropriate command-line version of Matlab for data preprocessing. On Dr. McKay's machine, this string is `/Applications/MATLAB_R2019a.app/bin/matlab -nojvm -nodesktop -nosplash -r`.
1. "knit" `R/msBalance.Rmd` to produce `R/msBalance.html`. This will call matlab to create intermediary files if they are missing.

## Notes related to the fact that there is a mixture of Matlab and R code
1. The source files are in `.mat` format.
1. There is a mixture of matlab preprocessing code (files in `m/`) into `.csv` files that R can then read (files in `R/`).
1. A directory `csv/` is created to hold these files. The contents of this directory are not automatically loaded to GitHub.

## How to download from GitHub for development
1. In RStudio, start a new Project: `File > New Project > Version Control > Git`.
1. In the “repository URL” paste: https://github.com/jlucasmckay/msBalance.
1. On OS X, you can make this under `tmp` - this is a temporary file that can be deleted later.
1. If there are files that need to be added to the remote, they can be placed in this folder or dragged / dropped into the GitHub interface.
1. Stage changed files, commit, and push changes with the git panel.