### 2020-01-17 R/msBalance.Rmd
updated table and figure captions

---
title: "Balance responses in MS"
author:
- J. Lucas McKay
- Daniel S. Peterson
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
bibliography: bibliography.bib
csl: nature.csl
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
# set path variables
source("../paths.m")

# if csv files do not exist, run matlab batch steps
if (!file.exists("../csv/c.csv")|!file.exists("../csv/m.csv")){
  system(paste(matlabCommand, "'cd ../m; loadData; exit'"))
}
```

# Introduction
This document describes the analysis of standing balance responses in people with and without multiple sclerosis (MS) using wavelet-based functional ANOVA (wfANOVA[@mckay2013]).
MS has profound effects on balance,[@vanliew2019] including altering protective stepping behaviors.[@peterson2016]
The main objective of this study was to examine the impact of MS on the pattern of magnitude and timing of muscle activation during feet-in-place balance responses.

# Methods

```{r loadData}
library(vroom)
library(dplyr)
library(magrittr)
library(stringr)
library(lubridate)
`%contains%` = function(s, pattern) grepl(pattern, s)
`%omits%` = function(s, pattern) !grepl(pattern, s)

# load participant-level information
m = vroom("../csv/m.csv") %>% mutate(Group = "MS") %>% mutate(Age = round(age)) %>% mutate(Sex = toupper(gender)) %>% rename(Participant = Subj)
c = vroom("../csv/c.csv") %>% mutate(Group = "Control") %>% mutate(Age = round(age)) %>% mutate(Sex = toupper(gender)) %>% rename(Participant = Subj)

# concatenate the participants. note that some of them appear to be coded as follow-ups. this is very useful, but may not be needed for the first report. they appear to be coded as such by the presence of "F1", "F2" in the participant name. convert this to a visit number (1, 2, 3, 4)
participants = full_join( m %>% select(-trial, -age, -gender) %>% distinct(), c %>% select(-trial, -age, -gender) %>% distinct() ) %>%
  mutate(`Visit Number` = ifelse(Participant %omits% "F\\d", 1, as.numeric(str_extract(str_extract(Participant, "F\\d"),"\\d"))+1)) %>% 
  mutate(`Visit Date` = date(dmy_hms(obsDate))) %>% select(-obsDate) %>% 
  mutate(`Visit ID` = Participant) %>% 
  mutate(`Visit Code` = str_extract(Participant, "F\\d")) %>% 
  mutate(Participant = str_replace(Participant, "F\\d", "")) %>% 
  mutate(Group = factor(Group, levels = c("MS","Control")))

# identify enrollment date for each participant and calculate amount of observation time
participants = participants %>% left_join(participants %>% group_by(Participant) %>% summarize(`Enrollment Date` = min(`Visit Date`, na.rm = TRUE))) %>% 
  mutate(`Time Since Enrollment (Days)` = ifelse(Group == "MS", `Visit Date` - `Enrollment Date`, 0))

# calculate amount of follow-up
followUp = participants %>% select(Group, Participant, `Time Since Enrollment (Days)`) %>% group_by(Group, Participant) %>% summarize(`Study Visits Available (n)` = n(), `Follow-Up Available (Days)` = max(`Time Since Enrollment (Days)`))

# consider only baseline data, and get rid of participants with missing minimal demographic data
participants %<>% filter(!`Visit ID` %contains% "F") %>% filter(!is.na(Age)) %>% filter(Sex %in% c("M","F")) %>% filter(Age>0)

# add a group condition - note that if_else in dplyr and ifelse in base r are different!
participants %<>% mutate(Group = if_else(Participant %contains% "MSC", "Control", "MS")) %>%
  mutate(Group = relevel(factor(Group), "MS")) %>% select(Group, Participant, Age, Sex, everything())
```

```{r layout}
# layout tables and figures
library(captioner)
tableNumbers <- captioner(prefix = "Table")
captionDemographics = tableNumbers(name = "demographics", caption = "Clinical and demographic features of the study population.")
figureNumbers <- captioner(prefix = "Figure")
captionCartoon = figureNumbers(name = "cartoon", caption = "Perturbation directions used in testing protocol.")
```

## Data sources
This study analyzed existing electromyographic recordings during balance responses in people with and without MS.
An earlier analysis of a subset of these records is available.[@peterson2016]
Inclusion criteria for MS participants were as follows: diagnosis of MS by a MS neurologist and no relapses within 60 days.
Inclusion criteria for all participants included ability to walk 25 ft (7.6 m) without an assistive device, and no other known biomechanical or neurologic conditions affecting stepping or balance.
Clinical and demographic characteristics of the study sample are summarized in `r tableNumbers("demographics", display = "cite")`.

```{r demographics}
library(janitor)
library(kableExtra)
participants %>% tabyl(Sex, Group) %>% mutate(Total = MS + Control) %>% 
  kable(caption = captionDemographics) %>% kable_styling(full_width = F, position = "left")
```

## Assessment protocol
Complete details of the assessment protocol are avaialble in an earlier report.[@huisinga2014] 
Briefly, subjects stood on a set of 2 custom hydraulic force platforms that translated either forward or backward (`r figureNumbers("cartoon", display = "cite")`).
In this study, "perturbation direction" refers to the direction of postural sway caused by the surface translation.
Therefore, a backward perturbation refers to forward movement of the support surface causing backward body sway and the activation of tibialis anterior (TA),
whereas forward perturbations refer to backward translation of the support surface causing forward body sway and the activation of medial gastrocnemius (MG).
Subjects stood with arms folded across the chest, eyes open, with their feet at a fixed heel-to-heel distance of 10 cm.
Peak platform displacement was 4 cm, at a rate of 15 cm/s, which elicited feet-in-place balance corrections without stepping.
Surface electromyography was recorded from the right TA of all subjects with two 2.5-cm^2^ surface electrodes placed approximately 2 cm apart, with a ground electrode on the lateral condyle.

```{r perturbationFigure, out.width = "25%", fig.align = "left", fig.cap = captionCartoon}
knitr::include_graphics("figures/Perturbations.svg")
```

## Data processing
Each individual EMG record was registered to the onset of platform motion and resampled.

## Demographic and clinical characteristics
Demographic and clinical characteristics of the study population are summarized in `r tableNumbers("demographics", display = "cite")`.

***
### Funding sources and conflict of interest
This study was supported in part by National Institutes of Health grant K25HD086276.
The authors report no conflict of interest related to this study.

### Financial disclosures for the previous 12 months
Dr. McKay has received research funding or support from the National Institutes of Health. Dr. McKay has received consulting fees from BioCircuit Technologies.

***
# References